Marginally-Stable Thermal Equilibria of Rayleigh-Bénard Convection

### 2021-05-27 publication_materials/ms_rbc.tex
Kyle's suggestions reflected

\begin{abstract}
    Natural convection is ubiquitous throughout the physical sciences and engineering, yet many of its important properties remain elusive.
    In this investigation, we derive and solve a quasilinear form of the Rayleigh-Bénard problem by representing the perturbations in terms of marginally stable eigenmodes.
    The amplitude of each eigenmode is determined by requiring that the background state maintains marginal stability.
    The background temperature profile evolves due to the advective flux of each eigenmode, as well as diffusion.
    The entire calculation is one-dimensional, and can be run on a workstation.
    We find the background temperature field evolves to an equilibrium state, where the advective flux from the marginally-stable eigenmodes and the diffusive flux sum to a constant.
    These marginally-stable thermal equilibria (MSTE) are exact solutions of the quasilinear equations.
    The mean temperature profile has thinner boundary layers and larger Nusselt numbers than thermally-equilibrated 2D and 3D simulations of the full nonlinear equations.
    We find the Nusselt number scales like $\Nu \sim\Ra^{1/3}$.
    When an MSTE is used as initial conditions for a 2D simulation, we find that Nu quickly equilibrates without the burst of turbulence often induced by purely conductive initial conditions, but we also find that the kinetic energy is too large and viscously attenuates on a long viscous time scale.
\end{abstract}

@@

Absent solid evidence from direct numerical simulation, other methods have been developed to try to infer large-$\Ra$ behavior or otherwise gain insight.
In the presence of other physical effects (e.g., rotation, magnetic fields), one can sometimes derive an asymptotically consistent set of reduced equations \cite{Julien2007, Julien2012}.
Reduced models are useful because they allow us to study the problem with less expensive computations.
Another approach relates to unstable exact coherent states (ECS) which are steady solutions to the full nonlinear problem \cite{Waleffe, Sondak, Wen, chini_cells}. 
Simulations and analyses performed by \cite{Yalniz, Cvitanovic} suggest that chaotic solution trajectories might ``visit'' these ECS.
Should that be the case, it is crucial that we discover and classify such equilibria.

@@

In this paper employ the same strategy to the more complicated Rayleigh-B\'{e}nard convection problem.
In section~\ref{sec:model} we recall the underlying equations, and in section~\ref{sec:evolution} we describe how we evolve the background temperature profile while maintaining marginal stability.
Section~\ref{sec:properties} describes the properties of the marginally-stable thermal equilibria, in particular how the Nusselt number and characteristic wavenumbers vary with the Rayleigh number.
Finally, we describe the results of simulations initialized with marginally-stable thermal equilibria in section~\ref{sec:sims}, and conclude in section~\ref{sec:Discussion}.

\section{Model Setup}\label{sec:model}
We begin with the Boussinesq approximation for Rayleigh-Bénard Convection, nondimensionalized on the freefall time scale. 
The domain $\mathcal{D}$ is 2-dimensional, rectangular, and horizontally periodic with cartesian spatial coordinates $0 \leq x < 4$ and $-1/2 < z < 1/2$. 
The fluid of interest is constrained between two flat boundaries at $z = -1/2$ and $z = 1/2$ with fixed temperatures $1/2$ and $-1/2$ respectively. 
At both boundaries we specify impenetrable, no-slip conditions, such that the velocity $\vec{u} = u \hat{x} + w \hat{z} = \vec{0}$ at $z = \pm 1/2$, where $\hat{x}, \hat{z}$ are the unit vectors in the $x$ and $z$ directions. 
The equations of motion are then given by
\begin{align}
    \nabla \cdot \vec{u} &= 0 \label{EQ:motion1}\\
    \frac{\partial \vec{u}}{\partial t} + \vec{u} \cdot \nabla \vec{u} &= - \nabla p + T \hat{z} + \mathcal{R} \nabla^2 \vec{u} \label{EQ:motion2}\\
    \frac{\partial T}{\partial t} + \vec{u} \cdot \nabla T &= \mathcal{P} \nabla^2 T \label{EQ:motion3}
\end{align}
where $p$ is pressure and $T$ is temperature. 
%For completeness, we specify a final boundary condition $p = p_0$ at $z = - 1/2$. 
Any system of this form can be characterized by its dimensionless Rayleigh number $\Ra = \frac{g\alpha L^3 \Delta T}{\nu \kappa}$ and Prandtl number $\Pr = \frac{\nu}{\kappa}$, where $g, \, \alpha, \, L_z, \, \Delta T, \nu, \kappa$ are the gravitational acceleration, coefficient of thermal expansion, domain height, imposed temperature difference, kinematic viscosity, and thermal diffusivity respectively. 
In this paper, we fix $\Pr = 1$.
For convenience, we define
\begin{equation}
\mathcal{R} = \sqrt{\frac{\Pr}{\Ra}}, \qquad \mathcal{P} = \frac{1}{\sqrt{\Pr \Ra}}.
\end{equation}

To derive the quasilinear form, we posit that an arbitrary field $f$ can be represented as the sum of a mean profile (denoted by $\overline{f}$) and a perturbation function (denoted by $f'$).
\begin{align}
    \vec{u}(x, z, t) &= \vec{u'}(x, z, t) \label{EQ:reynolds_dc_u}\\
    &= u'(x, z, t)\hat{x} + w'(x, z, t)\hat{z} \\
    T(x, z, t) &= \overline{T}(z, t) + T'(x, z, t) \label{EQ:reynolds_dc_T}\\
    p(x, z, t) &= \overline{p}(z, t) +  p'(x, z, t) \label{EQ:reynolds_dc_p}.
\end{align}
where the mean-velocity components vanish due to incompressibility and symmetry. Perturbations are defined to have no horizontal-average
\begin{equation}
    \langle f'(x, z, t) \rangle_x \equiv \int_{0}^4 f'(x, z, t) dx = 0.
\end{equation}
%Assuming the existance of nontrivial solutions, we also fix the volume-average over the entire domain
%\begin{equation}
%  \langle T'^2 \rangle_{\mathcal{D}} = 1
%\end{equation}
%thereby alleviating any ambiguity in $A(t)$. 
Substituting \eq{EQ:reynolds_dc_T} into \eq{EQ:motion3} and taking the horizontal-average reduces the system to a simple initial value problem (IVP) for $\overline{T}$
\begin{equation}
  \frac{\partial \overline{T}}{\partial t} + \frac{\partial}{\partial z} \langle w'T' \rangle_x = \mathcal{P} \frac{\partial^2 \overline{T}}{\partial z^2}, \label{EQ:T0_IVP}
\end{equation}
with associated boundary conditions $\overline{T}(-1/2, t) = 1/2$ and $\overline{T}(1/2, t) = -1/2$. It should be noted that we could obtain a similar IVP for $u$ by breaking symmetry and considering some nontrivial mean horizontal flow $\overline{u}(z, t)$. However we must have $\overline{w}(z, t) = 0$ due to incompressibility.

Substituting \eq{EQ:reynolds_dc_p} into \eq{EQ:motion2} and taking the horizontal average reveals that the mean pressure field $\overline{p}(z)$ must satisfy
\begin{equation}
    0 = -\frac{\partial \overline{p}}{\partial z} + \overline{T}\label{EQ:p_bar}
\end{equation}

@@

For each $k_x$, we can assess the stability of the perturbations by solving for the eigenvalue $s$, whose imaginary component $\sigma$ plays the role of an exponential growth rate. 
Positive eigenvalues indicate that the system is unstable to small disturbances of wavenumber $k_x$, while negative eigenvalues indicate stability. 
A complete linear stability analysis requires solution over the full spectrum of wavenumbers. 
The prototypical case is used to demonstrate that the critical Rayleigh number $\Ra_c = 1708$ when $\frac{\partial \overline{T}}{\partial z} = -1$.

@@

Consider a marginally-stable temperature profile $\overline{T}(z, t_0)$.
By definition, its maximum growth rate is 0.
Diffusing $\overline{T}(z, t_0)$ tends to increase its growth rates while ignoring the diffusive term and evolving according to advection tends to decrease its growth rates.
$A$ must be selected such that these two influences are equal and opposite.
We can measure the effects of diffusion and advection on the maximum eigenvalue by solving two new IVPs
\begin{align}\label{eqn:diffusion_only}
    &\frac{\partial \overline{T}_{\rm{diff}}}{\partial t} = \mathcal{P} \frac{\partial^2 \overline{T}_{\rm{diff}}}{\partial z^2} \\
    &\frac{\partial \overline{T}_{\rm{adv}}}{\partial t} + \frac{\partial}{\partial z} \langle w'T' \rangle_x = 0 
\end{align}
where $\overline{T}_{\rm{diff}}$ and $\overline{T}_{\rm{adv}}$ denote to the diffused and advected temperature profiles.
We initialize $\overline{T}_{\rm{diff}}$ and $\overline{T}_{\rm{adv}}$ with $\overline{T}(z, t_0)$, and evolve the equations to $t=t_0+\Delta t$.
Suppose $k_0$ is the wavenumber of the marginally-stable mode.
Then, using the diffused and advected profiles, we compute the eigenvalues $\sigma_{\rm{diff}}$ and $\sigma_{\rm{adv}}$ at $k_x = k_0$ and $t=t_0+\Delta t$.
A preliminary approximation $\Lambda^2_0$ of the appropriate amplitude $A^2$ is given by
\begin{equation}
    A^2 \approx \Lambda^2_0 \equiv -\frac{\sigma_{\rm{diff}}}{\sigma_{\rm{adv}}} \label{EQ:amp_approx}
\end{equation}
where the square in $A^2$ and $\Lambda^2_0$ is due to the quadratic nonlinearity of the eigenfunctions in the advective term and the minus sign is due to the fact that $\sigma_{\rm{adv}} < 0$ and $\sigma_{\rm{diff}} > 0$ in general.
We illustrate these trends in Figure~\ref{fig:iteration_spectra}. 
The subscript 0 denotes a preliminary estimate which we later refine.

@@

\subsection{Treatment of Multiple Marginally-Stable Modes} \label{sec:multiple_modes}
In most cases, we encounter eigenvalue spectra with multiple marginal modes.
To accommodate this we generalize the advective term in \eq{EQ:T0_IVP} to accommodate $N$ simultaneously marginal modes
\begin{equation}
    \langle w' T' \rangle_x = \sum_{n = 1}^{N} 2 A_n^2  \Re\left[ W_n \theta_n^* \right] 
\end{equation}
where $W_n$ and $\theta_n$ are the marginally stable eigenmodes associated with different values of $k_x$. 
There are now $N$ modes, each with their own amplitude to solve for and eigenvalue to keep marginally-stable. 
Given a small fixed time step $\Delta t$, let $\vec{A^2}$ be the amplitude vector and $\vec{\sigma}(\vec{A^2})$ be the eigenvalues of the $N$ modes.
We are searching for $\vec{\tilde{A}^2}$ that satisfies of $\vec{\sigma}(\vec{\tilde{A}^2})=0$.
We approximate the amplitudes with $\vec{\Lambda^2_0}$ by generalizing \eq{EQ:amp_approx}:
\begin{equation}
    \vec{\tilde{A}^2} \approx \vec{\Lambda^2_0} = -\Sigma_{\rm{adv}}^{-1} \vec{\sigma_{\rm{diff}}}.
    \label{EQ:AN_approx}
\end{equation}
Here $\vec{\sigma_{\rm{diff}}} = \vec{\sigma} (\vec{0})$ refers to the eigenvalues of the background temperature after evolving under only diffusion (equation \ref{eqn:diffusion_only}) for a time interval $\Delta t$.
We account for the influence of $N$ advection terms on $N$ eigenvalues (one for each mode) by constructing an eigenvalue matrix $\Sigma_{\rm{adv}}$.
To calculate element of $\Sigma_{\rm{adv}}$ at row $i$ and column $j$, we evolve the background temperature under only advection by mode $j$ for a time interval $\Delta t$.
Then the $i,j$ component of $\Sigma_{\rm{adv}}$ is given by the eigenvalue of the $i$th mode.

@@

Figure~\ref{fig:T0_profiles} gives temperature profiles for $\Ra = 10^8$, where the initial profile is outlined in appendix~\ref{sec:initial_profile}. 
We ran a direct numerical simulation (DNS) of \eqss{EQ:motion1}{EQ:motion3} with \texttt{Dedalus}, and also plot the horizontal- and time- averaged temperature profile. 
The temperature boundary layers in the DNS are wider than in the MSTE, which is in turn wider than the boundary layer of the initial temperature profile.
Performing an eigenvalue solve by setting $\overline{T}$ equal to the DNS profile yields unstable eigenvalues. 
This suggests that MSTE might maximize boundary layer thickness, subject to the marginal stability constraint.

\begin{figure*}
    \centering
    \begin{tabular}{@{}c@{}}
        \includegraphics[width=3.4in]{flux_sup_n.pdf}
    \end{tabular}
    \begin{tabular}{@{}c@{}}
        \includegraphics[width=3.4in]{EV_spectra_2ra.pdf}
    \end{tabular}
    \caption{Heat fluxes (left) and eigenvalue spectra (right) of equilibrated states with $\Ra = 2 \times 10^5$ (top) and $\Ra = 10^9$ (bottom). 
    The heat flux profiles are symmetric about $z=0$, so we only plot for $z<0$.
    We plot the advection profiles for marginally-stable modes. 
    At low $\Ra$, a single mode with $k_x = 1.5\pi$ is sufficient to oppose boundary layer diffusion and facilitate heat flux throughout the bulk of the domain. 
    At large $\Ra$, high-wavenumber modes contribute pronounced small-scale advection profiles which tightly hug the thin boundary layers. 
    A combination of of progressively wider advection profiles is necessary to transition to bulk advection provided by the $k_x = 1.5\pi$ mode.}
    \label{fig:flux}
\end{figure*}

The most resilient and unexpected feature of MSTE temperature profiles are the pronounced dips adjacent to the boundary layers. 
These dips appear in every solution, regardless of $\Ra$. 
Physically, they correspond to thin layers in which the mean temperature gradient reverses, contradicting an important hypothesis of \cite{Malkus_1954,Howard_1966}. 
This counter-diffusion, which opposes overall heat transfer, is overcome by the coinciding advective flux, shown in Figure~\ref{fig:flux}. 
We do not understand the source of these dips, but similar temperature gradient reversals were reported by \cite{chini_cells} along the midlines of 2D convective cellular solutions at $\Ra \sim 10^6$.
In that case, the reversals were due to nonlinear advection, which is not present in our quasilinear model.

In Figure~\ref{fig:flux}, we give heat flux profiles and eigenvalue spectra for two cases: $\Ra = 2 \times 10^5$ (top) and $\Ra = 10^9$ (bottom). 
For $\Ra = 2 \times 10^5$, there is a single marginal mode at $k_x = 1.5\pi$ whose advective flux occupies the bulk of the domain. 
These states have wide boundary layers, with diffusive fluxes which gradually subside as advection becomes the dominant flux component. 
Transitional regions occur over a smaller length scale for $\Ra = 10^9$ where the shift from diffusion to advection is sharp.
At $\Ra=10^9$ we find five marginally-stable modes are necessary to reach an MSTE.
Thin advection profiles, belonging to high-wavenumber modes with $k_x=23.5\pi, \, 24\pi$, hug the boundary layer. 
Closer to the bulk of the domain, we see wider advection profiles corresponding to modes in a second group of marginal modes $k_x = 6\pi, \, 6.5\pi$.
The $k_x = 1.5\pi$ mode forms the large-scale convective cell structures observed in DNS, again occupying the bulk of the domain.
The pairs of modes $k_x = 6\pi, \, 6.5\pi$ and $k_x=23.5\pi, \, 24\pi$ are each associated with a single maximum in our plots of growth rate $\sigma$ as a function of $k_x$ (lower right panel of Figure~\ref{fig:flux}).
If we allowed wavenumbers to vary continuously, there would be an unstable mode between these pairs of wavenumbers.
However, since we have fixed the horizontal size of our domain, we are left with pairs of discrete marginal modes.

@@

We find the MSTE can be characterized by their boundary layer height.
In Figure~\ref{fig:bl_ra}, we illustrate the scaling behavior of the boundary layer height $\delta = \Ra^{-1/3}$. 
This is consistent with Malkus' classical marginal-stability theory, a scaling argument which perceives the boundary regions as subdomains which are themselves marginally-stable \cite{Malkus_1954}.

@@

Using the mean temperature in a statistically-steady DNS as a background state for an eigenvalue problem yields positive eigenvalues: the system is in a perpetual state of instability. 
Unstable modes tend to stabilize the system rapidly, creating a negative feedback loop whose average state is linearly unstable. 
Future work could adjust the marginal-stability criterion to allow for moderately unstable modes.
Should the fast and slow time scales not be entirely separate, these modes might persist for long times. 
This might lead to greater agreement with DNS.

>>>>

\begin{abstract}
    Natural convection is ubiquitous throughout the physical sciences and engineering, yet many of its important properties remain elusive.
    To study convection in a novel context, we derive and solve a quasilinear form of the Rayleigh-Bénard problem by representing the perturbations in terms of marginally stable eigenmodes.
    The amplitude of each eigenmode is determined by requiring that the background state maintains marginal stability.
    The background temperature profile evolves due to the advective flux of every marginally-stable eigenmode, as well as diffusion.
    To ensure marginal stability and to obtain the eigenfunctions at each timestep, we perform a one-dimensional eigenvalue solve on each of the allowable wavenumbers.
    The background temperature field evolves to an equilibrium state, where the advective flux from the marginally-stable eigenmodes and the diffusive flux sum to a constant.
    These marginally-stable thermal equilibria (MSTE) are exact solutions of the quasilinear equations.
    The mean temperature profile has thinner boundary layers and larger Nusselt numbers than thermally-equilibrated 2D and 3D simulations of the full nonlinear equations.
    We find the Nusselt number scales like $\Nu \sim\Ra^{1/3}$.
    When an MSTE is used as initial conditions for a 2D simulation, we find that Nu quickly equilibrates without the burst of turbulence often induced by purely conductive initial conditions, but we also find that the kinetic energy is too large and viscously attenuates on a long viscous time scale.
    This is due to the thin temperature boundary layers which diffuse heat more effectively, thereby requiring high-velocity advective flows to reach an equilibrium.
\end{abstract}

@@

Absent solid evidence from direct numerical simulation, other methods have been developed to try to infer large-$\Ra$ behavior or otherwise gain insight.
In the presence of other physical effects (e.g., rotation, magnetic fields), one can sometimes derive an asymptotically consistent set of reduced equations \cite{Julien2007, Julien2012}.
Reduced models are useful because they allow us to study the problem with less expensive computations.
Another approach relates to unstable exact coherent states (ECS) which are steady solutions to the full nonlinear problem \cite{Waleffe, Sondak, Wen, chini_cells}. 
Simulations and analyses performed by \cite{Yalniz, Cvitanovic} suggest that chaotic solution trajectories might intermittently resemble these ECS.
Should that be the case, it is crucial that we discover and classify such equilibria.

@@

In this paper employ the same strategy to the more complicated Rayleigh-B\'{e}nard convection problem.
In section~\ref{sec:model} we recall the underlying equations, and in section~\ref{sec:evolution} we outline how we evolve the background temperature profile while maintaining marginal stability.
Section~\ref{sec:properties} pertains to the properties of the marginally-stable thermal equilibria, in particular how the Nusselt number and characteristic wavenumbers vary with the Rayleigh number.
Finally, we analyze the results of simulations initialized with marginally-stable thermal equilibria in section~\ref{sec:sims}, and conclude in section~\ref{sec:Discussion}.

\section{Model Setup}\label{sec:model}
We begin with the Boussinesq approximation for Rayleigh-Bénard Convection, nondimensionalized on the free-fall time scale. 
The domain $\mathcal{D}$ is 2-dimensional, rectangular, and horizontally periodic with cartesian spatial coordinates $0 \leq x < 4$ and $-1/2 < z < 1/2$.
We define the corresponding domain width $L_x = 4$ and height  $L_z = 1$.
The fluid of interest is constrained between two flat boundaries at $z = -1/2$ and $z = 1/2$ with fixed temperatures $1/2$ and $-1/2$ respectively. 
At both boundaries we specify impenetrable, no-slip conditions, such that the velocity $\vec{u} = u \hat{x} + w \hat{z} = \vec{0}$ at $z = \pm 1/2$, where $\hat{x}, \hat{z}$ are the unit vectors in the $x$ and $z$ directions. 
The equations of motion are then given by
\begin{align}
    \nabla \cdot \vec{u} &= 0 \label{EQ:motion1}\\
    \frac{\partial \vec{u}}{\partial t} + \vec{u} \cdot \nabla \vec{u} &= - \nabla p + T \hat{z} + \mathcal{R} \nabla^2 \vec{u} \label{EQ:motion2}\\
    \frac{\partial T}{\partial t} + \vec{u} \cdot \nabla T &= \mathcal{P} \nabla^2 T \label{EQ:motion3}
\end{align}
where $p$ is pressure and $T$ is temperature. 
%For completeness, we specify a final boundary condition $p = p_0$ at $z = - 1/2$. 
A Boussinesq convection system of this form can be characterized by its dimensionless Rayleigh number $\Ra = \frac{g\alpha L^3_z \Delta T}{\nu \kappa}$ and Prandtl number $\Pr = \frac{\nu}{\kappa}$, where $g, \, \alpha, \, L_z, \, \Delta T, \nu, \kappa$ are the gravitational acceleration, coefficient of thermal expansion, domain height, imposed temperature difference, kinematic viscosity, and thermal diffusivity respectively. 
In this paper, we fix $\Pr = 1$.
For convenience, we define
\begin{equation}
\mathcal{R} = \sqrt{\frac{\Pr}{\Ra}}, \qquad \mathcal{P} = \frac{1}{\sqrt{\Pr \Ra}}.
\end{equation}

To derive the quasilinear form, we posit that an arbitrary field $f$ can be represented as the sum of a mean profile (denoted by $\overline{f}$) and a perturbation function (denoted by $f'$).
\begin{align}
    \vec{u}(x, z, t) &= \vec{u'}(x, z, t) \label{EQ:reynolds_dc_u}\\
    &= u'(x, z, t)\hat{x} + w'(x, z, t)\hat{z} \\
    T(x, z, t) &= \overline{T}(z, t) + T'(x, z, t) \label{EQ:reynolds_dc_T}\\
    p(x, z, t) &= \overline{p}(z, t) +  p'(x, z, t) \label{EQ:reynolds_dc_p}.
\end{align}
where the mean-velocity components vanish due to incompressibility and symmetry. Perturbations are defined to have no horizontal-average
\begin{equation}
    \langle f'(x, z, t) \rangle_x \equiv \int_{0}^{L_x} f'(x, z, t) dx = 0.
\end{equation}
%Assuming the existance of nontrivial solutions, we also fix the volume-average over the entire domain
%\begin{equation}
%  \langle T'^2 \rangle_{\mathcal{D}} = 1
%\end{equation}
%thereby alleviating any ambiguity in $A(t)$. 
Substituting \eq{EQ:reynolds_dc_T} into \eq{EQ:motion3} and taking the horizontal-average reduces the system to a simple initial value problem (IVP) for $\overline{T}$
\begin{equation}
  \frac{\partial \overline{T}}{\partial t} + \frac{\partial}{\partial z} \langle w'T' \rangle_x = \mathcal{P} \frac{\partial^2 \overline{T}}{\partial z^2}, \label{EQ:T0_IVP}
\end{equation}
with associated boundary conditions $\overline{T}(-1/2, t) = 1/2$ and $\overline{T}(1/2, t) = -1/2$. It should be noted that we could obtain a similar IVP for $u$ by breaking reflection symmetry about the midplane $z = 0$ and considering some nontrivial mean horizontal flow $\overline{u}(z, t)$. However we must have $\overline{w}(z, t) = 0$ due to incompressibility.

Substituting \eq{EQ:reynolds_dc_p} into \eq{EQ:motion2} and taking the horizontal average reveals that the mean pressure field $\overline{p}(z, \, t)$ must satisfy
\begin{equation}
    0 = -\frac{\partial \overline{p}}{\partial z} + \overline{T}. \label{EQ:p_bar}
\end{equation}

@@

For each $k_x$, we can assess the stability of the perturbations by solving for the eigenvalue $s$, whose imaginary component $\sigma$ plays the role of an exponential growth rate. 
Positive eigenvalues indicate that the system is unstable to small disturbances whose Fourier decomposition includes a nontrivial component of wavenumber $k_x$.
Negative eigenvalues indicate stability. 
A complete linear stability analysis requires solution over the full spectrum of wavenumbers. 
The prototypical case is used to demonstrate that the critical Rayleigh number $\Ra_c = 1708$ when $\frac{\partial \overline{T}}{\partial z} = -1$.

@@

Consider a marginally-stable temperature profile $\overline{T}(z, t_0)$.
By definition, its maximum growth rate is 0.
Diffusing $\overline{T}(z, t_0)$ tends to increase its growth rates while ignoring the diffusive term and evolving according to advection tends to decrease its growth rates.
$A$ must be selected such that these two influences are equal and opposite.
We can measure the effects of diffusion and advection on the maximum eigenvalue by solving two new IVPs
\begin{align}\label{eqn:diffusion_only}
    &\frac{\partial \overline{T}_{\rm{diff}}}{\partial t} = \mathcal{P} \frac{\partial^2 \overline{T}_{\rm{diff}}}{\partial z^2} \\
    &\frac{\partial \overline{T}_{\rm{adv}}}{\partial t} + \frac{\partial}{\partial z} \langle w'T' \rangle_x = 0 
\end{align}
where $\overline{T}_{\rm{diff}}$ and $\overline{T}_{\rm{adv}}$ denote the diffused and advected temperature profiles.
We initialize $\overline{T}_{\rm{diff}}$ and $\overline{T}_{\rm{adv}}$ with $\overline{T}(z, t_0)$, and evolve the equations to $t=t_0+\Delta t$.
Suppose $k_0$ is the wavenumber of the marginally-stable mode.
Then, using the diffused and advected profiles, we compute the eigenvalues $\sigma_{\rm{diff}}$ and $\sigma_{\rm{adv}}$ at $k_x = k_0$ and $t=t_0+\Delta t$.
A preliminary approximation $\Lambda^2_0$ of the appropriate amplitude $A^2$ is given by
\begin{equation}
    A^2 \approx \Lambda^2_0 \equiv -\frac{\sigma_{\rm{diff}}}{\sigma_{\rm{adv}}} \label{EQ:amp_approx}
\end{equation}
where the square in $A^2$ and $\Lambda^2_0$ is due to the quadratic nonlinearity of the eigenfunctions in the advective term and the minus sign is due to the fact that $\sigma_{\rm{adv}} < 0$ and $\sigma_{\rm{diff}} > 0$ in general.
We illustrate these trends in Figure~\ref{fig:iteration_spectra}. 
The subscript 0 denotes a preliminary estimate which we later refine.
In general, when $\Lambda^2_0 > 0$, the root-finding algorithm is stable.
The alternative can occur when the timestep approaches machine precision, in which case $\sigma_{\rm{diff}}$ and $\sigma_{\rm{adv}}$ are subject to floating point errors.

@@

\subsection{Treatment of Multiple Marginally-Stable Modes} \label{sec:multiple_modes}
In most cases, we encounter eigenvalue spectra with multiple marginal modes.
To accommodate this we generalize the advective term in \eq{EQ:T0_IVP} to accommodate $N$ simultaneously marginal modes
\begin{equation}
    \langle w' T' \rangle_x = \sum_{n = 1}^{N} 2 A_n^2  \Re\left[ W_n \theta_n^* \right] 
\end{equation}
where $W_n$ and $\theta_n$ are the marginally stable eigenmodes associated with different values of $k_x$. 
The factor of 2 is due to horizontal averaging.
There are now $N$ modes, each with their own amplitude to solve for and eigenvalue to keep marginally-stable. 
Given a small fixed time step $\Delta t$, let $\vec{A^2}$ be the amplitude vector and $\vec{\sigma}(\vec{A^2})$ be the eigenvalues of the $N$ modes.
We are searching for $\vec{\tilde{A}^2}$ that satisfies of $\vec{\sigma}(\vec{\tilde{A}^2})=0$.
We approximate the amplitudes with $\vec{\Lambda^2_0}$ by generalizing \eq{EQ:amp_approx}:
\begin{equation}
    \vec{\tilde{A}^2} \approx \vec{\Lambda^2_0} = -\Sigma_{\rm{adv}}^{-1} \vec{\sigma_{\rm{diff}}}.
    \label{EQ:AN_approx}
\end{equation}
Here $\vec{\sigma_{\rm{diff}}} = \vec{\sigma} (\vec{0})$ refers to the eigenvalues of the background temperature after evolving under only diffusion (equation \ref{eqn:diffusion_only}) for a time interval $\Delta t$.
We account for the influence of $N$ advection terms on $N$ eigenvalues (one for each mode) by constructing an eigenvalue matrix $\Sigma_{\rm{adv}}$.
To calculate element of $\Sigma_{\rm{adv}}$ at row $i$ and column $j$, we evolve the background temperature under only advection by mode $j$ for a time interval $\Delta t$.
Then the $i,j$ component of $\Sigma_{\rm{adv}}$ is given by the eigenvalue of the $i$th mode.

@@

Figure~\ref{fig:T0_profiles} gives temperature profiles for $\Ra = 10^8$, where the initial profile is outlined in appendix~\ref{sec:initial_profile}. 
We ran a direct numerical simulation (DNS) of \eqss{EQ:motion1}{EQ:motion3} with \texttt{Dedalus}, and also plot the horizontal- and time- averaged temperature profile. 
The temperature boundary layers in the DNS are wider than in the MSTE, which is in turn wider than the boundary layer of the initial temperature profile.
Performing an eigenvalue solve by setting $\overline{T}$ equal to the DNS profile yields unstable eigenvalues. 
Every marginally-stable initial profile we experiment with ($\tanh$, $\rm{erf}$) has thinner boundary layers than the DNS and MSTE profiles.
In conjunction, these two facts suggest that MSTE maximize boundary layer thickness, subject to the marginal stability constraint.

\begin{figure*}
    \centering
    \begin{tabular}{@{}c@{}}
        \includegraphics[width=3.4in]{flux_sup_n.pdf}
    \end{tabular}
    \begin{tabular}{@{}c@{}}
        \includegraphics[width=3.4in]{EV_spectra_2ra.pdf}
    \end{tabular}
    \caption{Heat fluxes (left) and eigenvalue spectra (right) of equilibrated states with $\Ra = 2 \times 10^5$ (top) and $\Ra = 10^9$ (bottom). 
    The heat flux profiles are symmetric about $z=0$, so we only plot for $z<0$.
    We plot the advection profiles for marginally-stable modes. 
    At low $\Ra$, a single mode with $k_x = 1.5\pi$ is sufficient to oppose boundary layer diffusion and facilitate heat flux throughout the bulk of the domain. 
    At large $\Ra$, high-wavenumber modes contribute pronounced small-scale advection profiles which tightly hug the thin boundary layers. 
    A combination of of progressively wider advection profiles is necessary to transition to bulk advection provided by the $k_x = 1.5\pi$ mode. 
    This bulk mode mimics the large-scale convective cells which are characteristic of Rayleigh–Bénard convection, both in qualitative structure and number of cells in the domain.
    The eigenfunctions which compose the advective profiles are illustrated in Appendix~\ref{sec:eigenfunctions}.}
    \label{fig:flux}
\end{figure*}

The most resilient and unexpected feature of MSTE temperature profiles are the pronounced dips adjacent to the boundary layers. 
These dips appear in every solution, regardless of $\Ra$. 
Physically, they correspond to thin layers in which the mean temperature gradient reverses, contradicting an important hypothesis of \cite{Malkus_1954}. 
This counter-diffusion, which opposes overall heat transfer, is overcome by the coinciding advective flux, shown in Figure~\ref{fig:flux}. 
We do not understand the source of these dips, but similar temperature gradient reversals were reported by \cite{chini_cells} along the midlines of 2D convective cellular solutions at $\Ra \sim 10^6$.
In that case, the reversals were due to nonlinear advection, which is not present in our quasilinear model.

In Figure~\ref{fig:flux}, we give heat flux profiles and eigenvalue spectra for two cases: $\Ra = 2 \times 10^5$ (top) and $\Ra = 10^9$ (bottom). 
For $\Ra = 2 \times 10^5$, there is a single marginal mode at $k_x = 1.5\pi$ whose advective flux occupies the bulk of the domain. 
These states have wide boundary layers, with diffusive fluxes which gradually subside as advection becomes the dominant flux component. 
Transitional regions occur over a smaller length scale for $\Ra = 10^9$ where the shift from diffusion to advection is sharp.
At $\Ra=10^9$ we find five marginally-stable modes are necessary to reach an MSTE.
Thin advection profiles, belonging to high-wavenumber modes with $k_x=23.5\pi, \, 24\pi$, hug the boundary layer. 
Closer to the bulk of the domain, we see wider advection profiles corresponding to modes in a second group of marginal modes $k_x = 6\pi, \, 6.5\pi$.
The $k_x = 1.5\pi$ mode forms the same arrangement and number of large-scale convective cells observed in DNS, again occupying the bulk of the domain.
The pairs of modes $k_x = 6\pi, \, 6.5\pi$ and $k_x=23.5\pi, \, 24\pi$ are each associated with a single maximum in our plots of growth rate $\sigma$ as a function of $k_x$ (lower right panel of Figure~\ref{fig:flux}).
If we allowed wavenumbers to vary continuously, there would be an unstable mode between these pairs of wavenumbers.
However, since we have fixed the horizontal size of our domain, we are left with pairs of discrete marginal modes.

@@

We find the MSTE can be characterized by their boundary layer height.
In Figure~\ref{fig:bl_ra}, we illustrate the scaling behavior of the boundary layer height $\delta = \Ra^{-1/3}$. 
This is consistent with Malkus' classical marginal stability theory, a scaling argument which perceives the boundary regions as subdomains which are themselves marginally-stable \cite{Malkus_1954}.

@@

Using the mean temperature in a statistically-steady DNS as a background state for an eigenvalue problem yields positive eigenvalues: the system is in a perpetual state of instability. 
Unstable modes tend to stabilize the system rapidly, creating a negative feedback loop whose average state is linearly unstable. 
Future work could adjust the marginal stability criterion to allow for moderately unstable modes.
Should the fast and slow time scales not be entirely separate, these modes might persist for long times. 
This might lead to greater agreement with DNS.
